Capon Beamforming and Moving Objects - An Analysis of Lateral Shift-Invariance

### 2014-03-05 HRUIofMovingObjects/jrnl_article.tex
Small final adjustments.

In \cite{Cox1973}, Cox presents in great detail how the amount of mismatch between the selected steering vector and the signal propagation vector impacts the beamformer output in simple situations. He also states the following important fact about the Capon beamformer [the $\vec{k}_3$ processor]: 
\begin{quote}
"\textit{... a $\vec{k}_3$ processor requires more closely spaced beams than a ... $\vec{k}_1$-processor in order to avoid serious signal suppression effects being introduced on signals arriving from directions between the beams}", 
\end{quote}
where $\vec{k}_1$ is the delay-and-sum (DAS) beamformer. In Fig.\,\ref{fig:das_capon_beams} we have plotted the two-way response for DAS and the Capon beamformer using (26) and (32) in \cite{Cox1973}. A 64-element linear array with $\lambda/2$ element spacing, where $\lambda$ is the wavelength, is scanned over a sector equal to four times the array resolution using narrow band excitation. Located in the farfield of the sector is an object moving parallel to the array surface. The plot shows the high precision scanned output when the object is located at plus and minus half the required beam spacing for DAS [See (\ref{eq:resolution})]. %No spatially white noise is present. Hence, the interference-noise matrix ($\sigma_0^2\mat{Q}$ in \cite{Cox1973}) is the identity. 
The effect of typical subarray averaging and diagonal loading have been added to the Capon beamformer output (subarray length of 32 and a diagonal loading factor of 0.01). We see that a DAS beamformer, using the required beam spacing, could yield a $1.9$ dB scalloping loss \cite{Harris1978} if the object is moving parallel to the array. However, if the Capon beamformer is used with the same beam density the loss could be as large as $27$ dB. This imaging system is then highly sensitive to small spatial shifts, either by the object or the array itself. A $1.9$ dB scalloping loss for the Capon beamforming is obtained for this specific example when the beam spacing or object distance is decreased with a factor of 25.

\begin{figure}[!t]
\centerline{
\includegraphics[width=\linewidth]{\img fig1_2way.eps}
}
\caption{Estimated drop in output power mid-way between receive beams for delay-and-sum and Capon beamforming (with subarray length of 32 and a diagonal loading factor of 0.01) applied in a farfield-narrowband setting. %The scanned output from both beamformers is plotted for two different time instances. For the first instances an object is located at an angle equal to half the required beam spacing for delay-and-sum, and for the latter the object is moved an angular distance corresponding to the system resolution. If the beam density was reduced to one beam per system resolution the maximum scalloping loss is found at zero degrees.
}
\label{fig:das_capon_beams}
\end{figure}

@@

Figure \ref{fig:das_shiftinvariant} presents such a LSV-plot for an imaging system using the DAS beamformer with a beam spacing equal to (\ref{eq:resolution}). Dotted vertical lines indicate position of transmit-receive beam pairs. The point scatterer response has been simulated using Field II \cite{Jensen1992, Jensen1996a} using a 96 element phased array with a 22 mm azimuth aperture, which was excited with a 2.5 MHz and 1.5 cycles long pulse. Transmit focus in azimuth and elevation were configured to be equal and the point scatterer was moved through this point. The transmit focus is where the ultrasound beam is at its narrowest, hence shift variance effects will appear at this location first. The beam density was calculated  based on (\ref{eq:resolution}) with the speed of sound equal to 1540. We see that the contours in Fig.\,\ref{fig:das_shiftinvariant} are approximately diagonal (low P-MAE) and that the amplitude variation among the maximum points is minimal (low A-MAE). We therefore conclude that the system is locally laterally shift-invariant. %In this, and the following LSV-plots based on simulations, the scan sequence always has a beam at zero degrees and the point scatterer is located in the center of this beam at one time instance.

@@

Figure \ref{fig:capon_oversampling_pr} shows the phase rotation method applied on the same scan sequence as in Fig.\,\ref{fig:capon} with four and sixteen times oversampling on receive respectively ($q$-factor). Note that the figure is zoomed compared with Fig.\,\ref{fig:capon}. As with oversampling on transmit there is some variation when the point moves in and out of focus. The maximum scalloping loss is therefore large. The extreme local variations are however reduced significantly. Comparing Fig.\,\ref{fig:capon_oversampling} with Fig.\,\ref{fig:capon_oversampling_pr}, the errors introduced by the phase rotation method are best seen in sidelobes.

@@

\begin{figure*}[!t]
	\centerline{
		\subfloat[]{
			\includegraphics[width=0.5\linewidth]{\img fig5a.eps}
		}
		\subfloat[]{
			\includegraphics[width=0.5\linewidth]{\img fig5b.eps}
		}
	}
	\caption{Capon beamforming as described in Fig.\,\ref{fig:capon} with oversampling based on phase rotation. The figure is zoomed compared with Fig.\,\ref{fig:capon} to better see the beam-to-beam variation. a) Four times oversampling ($q=4$). b) Sixteen times oversampling ($q=16$).}
	\label{fig:capon_oversampling_pr}
\end{figure*}

>>>>

In \cite{Cox1973}, Cox presents in great detail how the amount of mismatch between the selected steering vector and the signal propagation vector impacts the beamformer output in simple situations. He also states the following important fact about the Capon beamformer [the $\vec{k}_3$ processor]: 
\begin{quote}
"\textit{... a $\vec{k}_3$ processor requires more closely spaced beams than a ... $\vec{k}_1$-processor in order to avoid serious signal suppression effects being introduced on signals arriving from directions between the beams}", 
\end{quote}
where $\vec{k}_1$ is the delay-and-sum (DAS) beamformer. In Fig.\,\ref{fig:das_capon_beams} we have plotted the two-way response for DAS and the Capon beamformer using (26) and (32) in \cite{Cox1973}. A 64-element linear array with $\lambda/2$ element spacing, where $\lambda$ is the wavelength, is scanned over a sector equal to four times the array resolution using narrow band excitation. Located in the farfield of the sector is an object moving parallel to the array surface. The plot shows the high precision scanned output when the object is located at plus and minus half the required beam spacing for DAS [See (\ref{eq:resolution})]. %No spatially white noise is present. Hence, the interference-noise matrix ($\sigma_0^2\mat{Q}$ in \cite{Cox1973}) is the identity. 
The effect of typical subarray averaging and diagonal loading have been added to the Capon beamformer output [subarray length of 32 and a diagonal loading factor of 0.01]. We see that a DAS beamformer, using the required beam spacing, could yield a $1.9$ dB scalloping loss \cite{Harris1978} if the object is moving parallel to the array. However, if the Capon beamformer is used with the same beam density the loss could be as large as $27$ dB. This imaging system is then highly sensitive to small spatial shifts, either by the object or the array itself. A $1.9$ dB scalloping loss for the Capon beamforming is obtained for this specific example when the beam spacing or object distance is decreased with a factor of 25.

\begin{figure}[!t]
\centerline{
\includegraphics[width=\linewidth]{\img fig1_2way.eps}
}
\caption{Estimated drop in output power mid-way between receive beams for delay-and-sum and Capon beamforming [with subarray length of 32 and a diagonal loading factor of 0.01] applied in a farfield-narrowband setting. %The scanned output from both beamformers is plotted for two different time instances. For the first instances an object is located at an angle equal to half the required beam spacing for delay-and-sum, and for the latter the object is moved an angular distance corresponding to the system resolution. If the beam density was reduced to one beam per system resolution the maximum scalloping loss is found at zero degrees.
}
\label{fig:das_capon_beams}
\end{figure}

@@

Figure \ref{fig:das_shiftinvariant} presents such a LSV-plot for an imaging system using the DAS beamformer with a beam spacing equal to (\ref{eq:resolution}). Dotted vertical lines indicate position of transmit-receive beam pairs. The point scatterer response has been simulated using Field II \cite{Jensen1992, Jensen1996a} using a 96 element phased array with a 22 mm azimuth aperture, which was excited with a 2.5 MHz and 1.5 cycles long pulse. Transmit focus in azimuth and elevation were configured to be equal and the point scatterer was moved through this point. The transmit focus is where the ultrasound beam is at its narrowest, hence shift variance effects will appear at this location first. The beam density was calculated  based on (\ref{eq:resolution}) with a speed of sound equal to 1540. We see that the contours in Fig.\,\ref{fig:das_shiftinvariant} are approximately diagonal (low P-MAE) and that the amplitude variation among the maximum points is minimal (low A-MAE). We therefore conclude that the system is locally laterally shift-invariant. %In this, and the following LSV-plots based on simulations, the scan sequence always has a beam at zero degrees and the point scatterer is located in the center of this beam at one time instance.

@@

Figure \ref{fig:capon_oversampling_pr} shows the phase rotation method applied on the same scan sequence as in Fig.\,\ref{fig:capon} with four and sixteen times oversampling ($q$-factor) on receive respectively. Note that the figure is zoomed compared with Fig.\,\ref{fig:capon}. As with oversampling on transmit there is some variation when the point moves in and out of focus. The maximum scalloping loss is therefore large. The extreme local variations are however reduced significantly. Comparing Fig.\,\ref{fig:capon_oversampling} with Fig.\,\ref{fig:capon_oversampling_pr}, the errors introduced by the phase rotation method are best seen in sidelobes.

@@

\begin{figure*}[!t]
	\centerline{
		\subfloat[]{
			\includegraphics[width=0.5\linewidth]{\img fig5a.eps}
		}
		\subfloat[]{
			\includegraphics[width=0.5\linewidth]{\img fig5b.eps}
		}
	}
	\caption{Capon beamforming as described in Fig.\,\ref{fig:capon} with oversampling based on phase rotation. The figure is zoomed compared with Fig.\,\ref{fig:capon} to better see the beam-to-beam variation. a) Four times oversampling on receive ($q=4$). b) Sixteen times oversampling on receive ($q=16$).}
	\label{fig:capon_oversampling_pr}
\end{figure*}
